---
title: "Initial Korean covid 19"
author: "Louie A"
date: "6/21/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## METHODS
Today we are going to look at Sex of the patients and the result to see who died and who lived. WE are going to use Cat~Cat to help us look at the data.


As the graph shows the Males dying more often than the females are dying. It is about 50-70, 50% of males are alive while 40% are dead. There are more males dying than females.


## Graphical results - use ggplot commands to do this  

#### Plot barplots ggplot way

```{r}
dd2 <-  kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  As the graph shows the Males dying more often than the females are dying. It is about 50-70, 50% of males are alive while 40% are dead. There are more males dying than females.






  
## Numerical results

### Numerical Summaries  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
In the percentages there is a difference, about 12% of all people that are dead, 36% are female while 64%are male. in the second chart females Are ar 98.7 alive and males are at only 97. female 1.6 died, males 3%


## Inferential results 
## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Conclusions

